### 2022-03-18 paper/localpsf.tex
restructuring paper

Convolutions can be performed using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, which allows PC approximations to be used as a fast surrogates for $\Aop$. Often it is necessary to perform further linear algebra operations involving $\Aop$, such as solving linear systems with $\Aop$ as the coefficient operator. Conventionally, one performs these linear algebra operations using Krylov methods, replacing $\Aop$ with its PC approximation \cite{PRODCONVKRYLOV}. The Krylov method is accelerated because performing matrix vector products with the product-convolution approximation is typically faster than performing operator actions with $\Aop$. However, there are several problems with this approach. First, the number of Krylov iterations can be large, particularly when $\Aop$ is high-rank. Second, complicated geometries are rarely discretized using regular grids, so using the FFT to perform convolutions requires transferring discretized functions back and forth between the irregular mesh and a regular grid. Third, the computational cost grows with the number of convolution kernels used. 
%If a large number of convolution kernels are used, this approach can be slow.
%Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}. 
Thus, 
%after constructing the product-convolution approximation, 
it is desirable to convert the PC approximation to other matrix formats that are amenable to faster linear algebra operations. Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}, because operators that are amenable to PC approximation are often sparse in the wavelet domain. It is straightforward to show that convolution-product approximations (transposes of PC approximations) are equivalent to low rank approximation of the pseudo-differential symbol of the operator, so one could perform linear algebra operations involving PC approximations using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET}. Here, we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.

Since our end goal is constructing a H-matrix, we are freed from strict adherence to the PC format. This allows us to modify the PC approach to address well-known deficiencies. PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while our method is based on an assumption that we call ``normalized local mean displacement invariance,'' which is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}). Also, PC approximations have well-known issues near boundaries called boundary artifacts, which are caused by the boundary artifically ``cutting off'' impulse responses. One must extend the impulse responses outside of the domain where they are undefined, but
%. Various extension techniques have been used, such as extension by zero, and reflection across the boundary, yet 
no known extension technique is fully satisfactory \cite{BOUNDARYETENSION}. In our method we do not have an issue with boundary artifacts because we temporarily exclude impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined. In exchange for these benefits, our approximation can no longer be written as a weighted sum of convolution operators, so we cannot apply our approximation to vectors via the FFT. But this is fine because our approximation is an H-matrix, which can be applied to vectors even faster.

@@

Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we need to go through an elaborate process to compute impulse responses and interpolate them. Can we build the H-matrix directly? In short, no, not efficiently. The reason is that classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products. Our impulse response interpolation lets us quickly compute approximations to matrix entries, which allows us to use these classical H-matrix construction methods. There are newer matrix-free methods for H-matrix approximation based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}. While these matrix-free H-matrix construction methods are asymptotically scalable in theory, from a practical point of view the required number of matrix vector products is unreasonably large. Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make these methods more feasible in the future.

>>>>

Convolutions can be performed using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, which allows PC approximations to be used as a fast surrogates for $\Aop$. Often it is necessary to perform further linear algebra operations involving $\Aop$, such as solving linear systems with $\Aop$ as the coefficient operator. Conventionally, one performs these linear algebra operations using Krylov methods, replacing $\Aop$ with its PC approximation \cite{PRODCONVKRYLOV}. The Krylov method is accelerated because performing matrix vector products with the product-convolution approximation is typically faster than performing operator actions with $\Aop$. However, there are several problems with this approach. First, the number of Krylov iterations can be large, particularly when $\Aop$ is high-rank. Second, complicated geometries are rarely discretized using regular grids, so using the FFT to perform convolutions requires transferring discretized functions back and forth between the irregular mesh and a regular grid. Third, the computational cost grows with the number of convolution kernels used. 
%If a large number of convolution kernels are used, this approach can be slow.
%Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}. 
Thus, 
%after constructing the product-convolution approximation, 
it is desirable to convert the PC approximation to other matrix formats that are amenable to faster linear algebra operations. Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}, because operators that are amenable to PC approximation are often sparse in the wavelet domain. Convolution-product approximations (transposes of PC approximations) are equivalent to low rank approximation of the pseudo-differential symbol of the operator, so one could perform linear algebra operations involving PC approximations using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET}. Here, we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.

Since our end goal is constructing an H-matrix, we are freed from strict adherence to the PC format. This allows us to modify the PC approach to address well-known deficiencies. PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while our method is based on an assumption that we call ``normalized local mean displacement invariance,'' which is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}). PC approximations have well-known issues near boundaries called boundary artifacts, which are caused by the boundary artifically ``cutting off'' impulse responses. One must extend the impulse responses outside of the domain where they are undefined, but
%. Various extension techniques have been used, such as extension by zero, and reflection across the boundary, yet 
no known extension technique is fully satisfactory \cite{BOUNDARYETENSION}. In our method we do not have an issue with boundary artifacts because we temporarily exclude impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined. Our approximation cannot be written as a weighted sum of convolution operators, so we cannot apply our approximation to vectors via the FFT. This is fine because our approximation is an H-matrix, which can be applied to vectors even faster.

@@

Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we need to go through an elaborate process of computing impulse responses and interpolating them. 
%Can we build the H-matrix directly? In short, no, not efficiently. 
The reason is that classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products. Our impulse response interpolation lets us quickly compute approximations to matrix entries, which allows us to use these classical H-matrix construction methods. There are newer matrix-free methods for H-matrix approximation based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}. While these matrix-free H-matrix construction methods are asymptotically scalable in theory, from a practical point of view the required number of matrix vector products is unreasonably large. Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods more feasible in the future.
